COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-06-16 main_aaron.tex
Added aaron's working file

@@

Recent observations of circumstellar disks by ALMA have revealed a rich variety of substructure. Features 
such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the disks. In some cases, these gap features are argued to be an indicator for the 
presence of a giant planet, either embedded in the disk \citep{2015MNRAS.453L..73D} or orbiting externally. A giant perturber can influence the 
structure of the continuum emission in a number of ways. A misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}. Highly eccentric perturbers can produce even more complicated structures through secular interactions 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}. Mean motion resonances (MMRs) have been shown to open gaps as well 
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

Collisions between planetesimals are thought to be a significant source of dust for many young circumstellar disks 
\citep[see][]{2008ARA&A..46..339W}. An expanding debris cloud from a planetesimal collision was recently detected in the 
Fomalhut system \citep{2020PNAS..117.9712G}. Although some amount of primordial dust is likely still present during the planet formation process, 
ongoing collisions between small bodies will augment this and could be used to trace the underlying dynamical activity of the planetesimals. 
\citet{2013ApJ...777L..31D} showed that collisional dust generated near the gap formed by a giant planet in a planetesimal disk should produce a 
distinctly observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological differences in 
the dust emission could be used to determine whether the planet had a circular or eccentric orbit \citep{2016ApJ...820...29D}.

Substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues which 
could be used to constrain the properties of the planet.  The width of a resonance is set by both the mass of the perturbing planet, along with the 
unperturbed eccentricity of the planetesimals, which is in turn set by the eccentricity of the planet through secular forcing. Therefore, the structure of 
second generation dust produced near MMRs encodes information about both the mass and eccentricity of the planet.

@@

To further elucidate this subject, we use the tree based N-body code {\sc ChaNGa}
\citep{2008IEEEpds...ChaNGa, 2015AphCom..2..1} to follow the collisional evolution of a planetesimal disk under the gravitational 
influence of a Jupiter sized body. Because particle positions are sorted into a tree structure, neighbor finding and collision detection 
can be done quickly and efficiently. This considerably relaxes the constraints on resolution and integration time. With this toolset, we 
explore the collision rate structure of the planetesimal disk in the vicinity of mean motion resonances. In particular, we would like to 
determine: 1. whether MMRs leave a detectable signature in the collisionally-generated dust. 2. If these signatures can be used to determine or 
constrain the orbital properties of the perturbing planet.

@@

\begin{equation}\label{eq:fd_so}
	f_{d} (\alpha) = \frac{1}{8} \left[ -5(p+q) + 4(p+q)^{2} - 2 \alpha D + 4(p+q) \alpha D + \alpha^{2} D^{2} \right] b^{p+q}_{1/2},
\end{equation}

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk. From numerical tests with {\sc ChaNGa}, we have found that this time step size preserves 
the symplectic nature of the integration without being too computationally expensive. Although using $\Delta T$ = 0.01 yr keeps the 
integration symplectic, we find that a small amount of artificial precession is produced. This is most noticeable at the inner edge of the
planetesimal disk. We reduce the base time step size by an additional 
factor of 4 to the keep longitude of perihelia of planetesimals near the inner edge of the disk from drifting too far over the course of our 
integrations.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/xy.png}
    \caption{The positions of the remaining planetesimals at the end of the e1m2 (left), e2m2 (center) and e3m2 (right) simulations. The red dot 
    indicates the position of the giant planet and the dashed line points in the direction of the planet's longitude of perihelion. Non-axisymmetric gaps 
    are apparent near the locations of MMRs. At higher eccentricities, more resonances become visible. At the 2:1 MMR, gap features at $\theta$ = 0 
    and $\theta = \pi$ follow the giant planet in its orbit.\label{fig:xy}}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/ae.png}
    \caption{The semimajor axes and eccentricities of the remaining planetesimals are shown, with the locations of prominent resonances indicated 
    by the vertical dashed lines. Libration of the critical angle drives large variations in eccentricity, which produce spikes in the a-e plane. Between 
    the resonances, the nonzero eccentricity is due to secular forcing by the planet.\label{fig:ae}}
\end{center}
\end{figure}

>>>>

\newcommand{\ACBc}[1]{\textcolor{red}{#1}}

@@

Recent observations of circumstellar disks by ALMA have revealed a rich variety of substructure. Features 
such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the disks. In \sout{some cases}\ACBc{many gas-rich systems}, these gap features are argued \sout{to be an indicator for}\ACBc{to indicate} the
presence of a giant planet, either embedded in the disk \citep{2015MNRAS.453L..73D} or orbiting externally, \ACBc{although the details of gas-dust interactions can have strong effects on the resulting morphologies (Dong).}
\ACBc{In gas-poor/debris systems,} a giant perturber can \ACBc{also} influence the 
structure of the \ACBc{dust} continuum emission \sout{ in a number of ways}. \ACBc{For example,} a misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}, highly eccentric perturbers can produce \sout{even more complicated} structures through secular interactions 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}\ACBc{, and} mean motion resonances (MMRs) \sout{have been shown to open gaps as well }\ACBc{can open gaps}
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

Collisions between planetesimals are thought to be \sout{a significant}\ACBc{the principal} source of dust for \sout{many young circumstellar disks}{debris disks}. 
\citep[see][]{2008ARA&A..46..339W}.  Although some amount of primordial dust is likely still present during the \sout{planet formation process}\ACBc{early stages of debris disk evolution}, 
ongoing collisions between small bodies will augment this and could be used to trace the underlying dynamical activity of the planetesimals. 
\ACBc{This process was directly detected when the putative object Fomalhaut b was confirmed to be an} expanding debris cloud\ACBc{, most likely} from a planetesimal collision \sout{was recently detected in the 
Fomalhaut system} \citep{2020PNAS..117.9712G}.

\citet{2013ApJ...777L..31D} showed that collisional dust \ACBc{that is} generated near the gap \sout{formed}\ACBc{opened} by a giant planet in a planetesimal disk should produce a 
distinctly observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological differences in 
the dust emission could be used to determine whether the planet had a circular or eccentric orbit \citep{2016ApJ...820...29D}.
\ACBc{Moreover,} substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues \sout{which}\ACBc{that} 
could be used to constrain the properties of the planet.  The width of a resonance is set by both the mass of the perturbing planet, along with the 
unperturbed eccentricity of the planetesimals, which is in turn set by the eccentricity of the planet through secular forcing. Therefore, the structure of 
second generation dust produced near MMRs encodes information about both the mass and eccentricity of the planet.

@@

To further elucidate this subject, we use the tree based N-body code {\sc ChaNGa}
\citep{2008IEEEpds...ChaNGa, 2015AphCom..2..1} to follow the collisional evolution of a planetesimal disk under the gravitational 
influence of a Jupiter sized body. Because particle positions are sorted into a tree structure, neighbor finding and collision detection 
can be done quickly and efficiently. This considerably relaxes the constraints on resolution and integration time. With this toolset, we 
explore the collision rate structure of the planetesimal disk in the vicinity of mean motion resonances. In particular, we would like to 
determine \ACBc{(1)} whether MMRs leave a detectable signature in the collisionally-generated dust\ACBc{, and (2) whether} these signatures can be used to determine or 
constrain the orbital properties of the perturbing planet.

 \ACBc{[Feel free to omit this, particularly because it has a self citation, but I think it is important to highlight that your results could be relevant beyond debris systems]  While the focus here is on the evolution of planetesimal disks in general, with the most direct application being debris systems, we note that the results may be relevant for interpreting structures observed in young, gas-rich disks. For example, assuming giant planet formation requires first the presence of planetesimals, then systems such as HL Tau that have purported embedded giant planets, could have localized regions of dust enhancement due to the corresponding planetesimal collisions  \citep{2017ApJ...850..103B}.  The main difference between such cases and debris disks is that the dust in the former situation couples to gas. }

@@

\begin{eqnarray}\label{eq:fd_so}
	f_{d} (\alpha) &=& \frac{1}{8} \left[ -5(p+q) + 4(p+q)^{2} - 2 \alpha D \right. \\ \nonumber
	                      & + & \left. 4(p+q) \alpha D + \alpha^{2} D^{2} \right] b^{p+q}_{1/2},
\end{eqnarray}

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk. From numerical tests with {\sc ChaNGa}, we have found that this time step size preserves 
the symplectic nature of the integration without being too computationally expensive. Although using $\Delta T$ = 0.01 yr keeps the 
integration symplectic, we find that a small amount of artificial precession is produced. This is most noticeable at the inner edge of the
planetesimal disk. We reduce the base time step size by an additional 
factor of 4 to the keep longitude of perihelia of planetesimals near the inner edge of the disk from drifting too far over the course of our 
integrations. \ACBc{Define too far.}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/xy.png}
    \caption{\ACBc{fix aspect ratino so a circle is a circle.}  The positions of the remaining planetesimals at the end of the e1m2 (left), e2m2 (center) and e3m2 (right) simulations. The red dot 
    indicates the position of the giant planet and the dashed line points in the direction of the planet's longitude of perihelion. Non-axisymmetric gaps 
    are apparent near the locations of MMRs. At higher eccentricities, more resonances become visible. At the 2:1 MMR, gap features at $\theta$ = 0 
    and $\theta = \pi$ follow the giant planet in its orbit.\label{fig:xy}}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/ae.png}
    \caption{\ACBc{annotations need to be larger.}  The semimajor axes and eccentricities of the remaining planetesimals are shown, with the locations of prominent resonances indicated 
    by the vertical dashed lines. Libration of the critical angle drives large variations in eccentricity, which produce spikes in the a-e plane. Between 
    the resonances, the nonzero eccentricity is due to secular forcing by the planet.\label{fig:ae}}
\end{center}
\end{figure}
